---
title: "visualization_1"
author: "Amrutha Banda"
output: github_document
date: "2025-09-25"
---


```{r setup, include=FALSE}
library(tidyverse)
library(ggridges)
library(p8105.datasets)
```

import the weather data
```{r}
data("weather_df")
```

Making out first plot
```{r}
ggplot(weather_df, aes(x = tmin, y = tmax))
```


```{r}
ggplot(weather_df, mapping= aes(x = tmin, y = tmax)) + 
  geom_point()
```


creates the same scatterplot but included piping 
```{r}
weather_df |>
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point()
```

will be shown in environment 
```{r}
ggp_weather_scatterplot=
  weather_df |>
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point()

ggp_weather_scatterplot
```

check that some rows are missing 
```{r}
weather_df |> 
 filter(is.na(tmax))

```


## Fancier scatterplots 

```{r}
weather_df |> 
  ggplot(aes(x=tmin, y=tmax, color=name)) + 
  geom_point(alpha= 0.3) +
  geom_smooth(se= FALSE)

```


Where you define aesthetics can matter 
```{r}
weather_df |> 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point(aes(color=name, size= prcp), alpha = 0.3) +
  geom_smooth(se = FALSE) 

```

use faceting real quick 
ˆ
```{r}

weather_df |> 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point(aes(color=name), alpha=0.3, size=0.8) +
  geom_smooth(se = FALSE) +
  facet_grid(. ~ name)

weather_df |> 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point(aes(color=name), alpha=0.3, size=0.8) +
  geom_smooth(se = FALSE) +
  facet_grid(name ~. )

```

Let's make a somehwta more interesting scatterplot

```{r}

weather_df |> 
  ggplot(aes(x = tmin, y = tmax, color=name)) + 
  geom_point(aes(size=prcp), alpha=0.3) +
  geom_smooth(se = FALSE) +
  facet_grid(. ~ name)
  
```


Learning Assessment: Write a code chain that starts with weather_df; focuses only on Central Park, converts temperatures to Fahrenheit, makes a scatterplot of min vs. max temperature, and overlays a linear regression line (using options in geom_smooth())

LA plot: 

whenever we create something new we use `mutate`
`se= FALSE` removes the error band 

```{r}
weather_df |> 
  filter(name == "CentralPark_NY") |> 
  mutate(
    tmax_fahr= tmax * (9/5) +32, 
    tmin_fahr= tmin * (9/5) +32
  ) |> 
  ggplot(aes(x = tmin_fahr, y = tmax_fahr)) + 
  geom_point() +
  geom_smooth(se= FALSE, method= "lm")
```

## Small Things 

```{r}
weather_df |> 
  ggplot(aes(x = tmin, y=tmax, color=name, shape=name)) + 
# geom_point(alpha=0.3) +
  geom_smooth(se= FALSE)

weather_df |> 
  ggplot(aes(x = tmin, y=tmax, color=name, shape=name)) +
  geom_smooth(se= FALSE) +
  geom_point()
```


```{r}
weather_df |> 
  ggplot(aes(x=tmin, y=tmax)) +
  geom_hex()

```


```{r}

weather_df |> 
  ggplot(aes(x=tmin, y=tmax)) +
  geom_point(color="#098")
```

## Univariate plots 

```{r}
weather_df |> 
  ggplot(aes(x=tmin)) +
  geom_histogram(color= "white", fill= "red")

```


```{r}
weather_df |> 
  ggplot(aes(x=tmin, color=name)) +
  geom_histogram()

weather_df |> 
  ggplot(aes(x=tmin, fill=name)) +
  geom_histogram()


```

How would I fix this? Maybe facet?

```{r}
weather_df |> 
  ggplot(aes(x=tmin, fill=name)) +
  geom_histogram() +
  facet_grid(name ~ .)
```

Make a density plot?

```{r}
weather_df |> 
  ggplot(aes(x=tmin, fill=name)) +
  geom_density(alpha=0.2) 
```

```{r}
weather_df |> 
  ggplot(aes(x=name, y=tmin)) +
  geom_boxplot(aes(fill=name))
```


violin plots 

```{r}

weather_df |> 
  ggplot(aes(x=name, y=tmin, fill=name)) +
  geom_violin()

```

ridge plot

```{r}
weather_df |> 
  ggplot(aes(x=tmin, y=name)) +
  geom_density_ridges()
```


Learning Assessment: Make plots that compare precipitation across locations. Try a histogram, a density plot, a boxplot, a violin plot, and a ridgeplot; use aesthetic mappings to make your figure readable

```{r}
weather_df |> 
  ggplot(aes(x=prcp, fill=name)) +
  geom_histogram() +
  facet_grid(name ~ .)

```

```{r}
weather_df |> 
  filter(prcp > 5, prcp<1000) |> 
  ggplot(aes(x=prcp, fill=name)) +
  geom_density(alpha=0.3)
```

```{r}

ggp_weather_violin= 
  weather_df |> 
  ggplot(aes(x=name, y=tmin, fill=name)) +
  geom_violin()

ggp_weather_violin

ggsave("plots/violin_plot.pdf", ggp_weather_violin, 
       width=8, height=6)
```


## Saving and embedding plots

embedding plots
```{r, fig.width=6, fig.height=4}

ggp_weather_violin
```

